california-chinook-microhaps: Data and code repository for \"A multipurpose microhaplotype panel for genetic analysis of California Chinook salmon\

### 2025-03-26 README.Rmd
Thumbnails and links in README

# choose the one that you want here depending on if you are on the
# GitHub README or are making an index.html for the docs directory
hdir <- Dbase

@@

@@

The following RMarkdown documents should be evaluated in order.
The script `render-numbered-Rmds.R` will do that (except for `002-allele-frequencies.Rmd`)
when run in the top level
of this repository.
Some RMarkdown documents rely on
outputs from previous ones.  Some of these RMarkdown documents include
calls to Unix utilities, so might not run on non-Unix or non-Linux architectures.

Outputs (figures, tables, R data objects, etc) from each RMarkdown document
are written to the `outputs/XXX` directories.  Intermediate files
written during evaluation of each RMarkdown document are written
to the `intermediates/XXX` directories. To facilitate working between
the cluster and a desktop/laptop, some outputs are written to
the `stored_results/XXX` directories which are version controlled and
included in this repo.

Thumbnails of the figures and tables generated by each RMarkdown document appear
below.  Additionally, at the top of each section is a link to the
compiled (HTML-version) of the RMarkdown document on GitHub Pages.

## `r make_link("001-align-coho-genome.Rmd")`

The products of this analysis go into creation of some of the trees (`009`) and also
in the designation of the alleles as ancestral or not in some of the haplotype rasters (`006`).

## `r make_link("002-allele-frequencies.Rmd")`

Calculation of allele frequency differences between early- and late-running forms
across the genome, and around GREB1L, and the generation of plots.

[![](./docs/images_and_thumbs/allele-freqs-figure_thumb.png)][alle_freqs]

## `r make_link("003-extract-johnson-creek-variants.Rmd")`

Using the previously assembled Chinook salmon genome from @narumGenomicVariationUnderlying2018
to ascertain variation from that Upper Columbia River summer/spring-run fish
around the GREB1L region.  The products of this analysis go into some later trees and
haplotype rasters.

## `r make_link("004-prepare-haplotypes.Rmd")`

Using BEAGLE to impute genotypes and infer haplotypes in 5 Mb around GREB1L.
The outputs of this analysis are used for a number of downstream analyses.

## `r make_link("005-annotating-variants-near-greb1l.Rmd")`

Using snpEff to annotate variants and then analyzing a few of them.  This
provides data that goes into Table S1 ("Two near-perfectly associated, non-synonymous variants in the GREB1L gene"),
and is also used later for the haplotype raster plots.

[![](./docs/images_and_thumbs/table-s2-2-candidate-snps_thumb.png)][snpeff_res]

## `r make_link("006-haplo-raster-plots.Rmd")`

Compiling some auxiliary information (gene/exon inclusion, etc.) and
then making haplotype raster plots using functions from the
'ecaRbioinf' package.

[![](./docs/images_and_thumbs/moderate-zoom-haplo-raster-with-JC-fish_thumb.png)][haplo_raster]
[![](./docs/images_and_thumbs/read-depth-raster_thumb.png)][read_depth_raster]

>>>>

# choose the one that you want here depending on if you are on the
# GitHub README or are making an index.html for the docs directory
hdir <- GPbase

@@

# Repo Contents

The most important data in this repository are in:

- `data/SWFSC-chinook-reference-baseline.csv.gz`: The actual reference data set---genotypes
of all the fish in the baseline.
- `data/baseline_repository_meta_complete.csv.gz`: Metadata from all those reference
data set fish

The rest of the repo is involved in analyses of those data sets.  
Here is a an annotated graphical overview of the contents of the
repo made using `tree`.   Some file-series are omitted and replaced with
ellipses.

```sh
.
├── .Rprofile                                             # For initiating renv
├── .gitignore                                            # standard git ignore file
├── 001-map-of-samples.Rmd                                # 8 Rmarkdown documents (where the analyses happen)
├── 002-GSI_and_Fst.Rmd                                  
├── 003-ckmr-sim-evaluation.Rmd
├── 004-lfar-and-wrap-allele-freqs.Rmd
├── 005-plotting-genome-locations-of-markers.Rmd
├── 006-winter-run-vs-all-cv-alle-freqs.Rmd
├── 007-whoa-on-chinook.Rmd
├── 008-lfar-gwas-and-markers.Rmd
├── R                                                     # Directory of R code that gets sourced and reused in the notebooks
│   ├── colors.R
│   ├── get_top_assignments.R
│   ├── manhattan_plot_funcs.R
│   ├── month_sample_sparkline.R
│   ├── plot_self_assigment_table.R
│   ├── shift_legend.R
│   ├── table-as-figure.R
│   └── taf-prep.R
├── README.Rmd                                            # Source for README and docs/index
├── README.md                                             # Produced from README.Rmd
├── Snakefile                                             # Snakefile to run all R notebooks
├── SnakemakeSubDirs                                      # Copies of other repos used for cluster-based analyses
│   └── cga-late-fall-bam-assoc                           # LFAR GWAS using angsd code uses Snakemake
│       ├── .gitignore
│       ├── README.Rmd
│       ├── Snakefile                                     # This has all the info on how to run the LFAR/GWAS analysis
│       ├── cga-late-fall-bam-assoc.Rproj
│       ├── envs
│       │   ├── angsd.yaml
│       │   └── samtools.yaml
│       ├── inputs
│       │   ├── lf_v_both.bamlist
│       │   ├── lf_v_both.ybin
│       │   ├── lf_v_frhf.bamlist
│       │   ├── lf_v_frhf.ybin
│       │   ├── lf_v_sanjo.bamlist
│       │   ├── lf_v_sanjo.ybin
│       │   ├── region_files
│       │   │   ├── unplaced-001.txt
│       │   │   ├── unplaced-002.txt
│       │   │   ├── ...
│       │   │   ├── unplaced-027.txt
│       │   │   └── unplaced-028.txt
│       │   └── scaffold_groups.tsv
│       ├── scripts
│       │   ├── manhattan-plot.R                          # Functions for making manhattan plots
│       │   └── prep
│       │       └── make-scaff-beds.R
│       └── sedna                                         # cluster profile for Snakemake
│           ├── config.yaml
│           └── status-sacct.sh
├── california-chinook-microhaps.Rproj                    # Rstudio project file
├── data                                                  # Main data files
│   ├── SWFSC-chinook-reference-baseline.csv.gz           # The actual reference data set
│   └── baseline_repository_meta_complete.csv.gz          # Meta data for all fish in reference data set
├── development                                           # Folder of various things that we ended up not using
│   ├── genepop-conversion.Rmd
│   └── genepop-conversion.nb.html
├── docs                                                  # Directory of html files served up on GitHub Pages
│   ├── 001-map-of-samples.html
│   ├── 002-GSI_and_Fst.html
│   ├── 003-ckmr-sim-evaluation.html
│   ├── 004-lfar-and-wrap-allele-freqs.html
│   ├── 005-plotting-genome-locations-of-markers.html
│   ├── 006-winter-run-vs-all-cv-alle-freqs.html
│   ├── 007-whoa-on-chinook.html
│   └── 008-lfar-gwas-and-markers.html
├── envs                                                  # YAMLS specifying conda environments for snakemake rules
│   └── pandoc.yaml
├── extras                                                # Scripts and code used in cleaning/standardizing the baseline prior to distribution 
│   ├── .gitignore
│   ├── Applying-New-Names-To-Baseline
│   │   ├── .gitignore
│   │   ├── applying-new-names-to-baseline.Rmd
│   │   ├── subRoSA_baseline_rubias_JUN2024_w_mixture_complete-rosa_lfar_vgll3six6_added.csv
│   │   └── subRoSA_baseline_rubias_with_rosa_wrap_lfar_vgll3six6_and_mixture_28JUN2024.csv
│   ├── Renaming-amplicons
│   │   ├── .gitignore
│   │   ├── Amplicon-renaming.Rmd
│   │   ├── ChinookFullPanelLociTable_May2024Updated.xlsx
│   │   ├── Otsh_v2.0_chroms.tsv
│   │   └── add-species-id.Rmd
│   └── Reordering-forward-and-reverse-primers
│       ├── OLD-Calif-Chinook-Amplicon-Panel-Information.csv
│       └── primer-pair-reorder.R
├── inputs                                                 # Small inputs used in most R notebooks
│   ├── Calif-Chinook-Amplicon-Panel-Information.csv
│   ├── Otsh_v1.0_chrom_lengths.tsv
│   ├── Otsh_v2.0_chrom_lengths.tsv
│   ├── Rivers_OR
│   │   ├── Rivers_OR.dbf
│   │   ├── Rivers_OR.prj
│   │   ├── Rivers_OR.sbn
│   │   ├── Rivers_OR.sbx
│   │   ├── Rivers_OR.shp
│   │   └── Rivers_OR.shx
│   ├── map-notations.tsv
│   ├── reference-collection-names.csv
│   └── wrap
│       └── chrom_lengths.txt
├── renv                                                   # Folder of renv R package management stuff 
│   ├── .gitignore
│   ├── activate.R
│   └── settings.json
├── renv.lock                                              # renv R package management lockfile
├── scripts                                                # Scripts used when running Snakemake or prepping the Snakefile
│   ├── R_input_and_output_lists_to_snakemake_skeleton.sh
│   └── render-rmd-for-snakemake.R
├── stored_results                                         # Results produced on the HPCC stored here for further analysis
│   ├── 5Mb_Chr34-cv-fall.mafs.gz
│   ├── 5Mb_Chr34-late-fall.mafs.gz
│   ├── lfar-gwas
│   │   ├── lf_v_frhf-NC_037130.1.lrt0.gz
│   │   └── lf_v_sanjo-NC_037130.1.lrt0.gz
│   ├── stored_p_vals.rds
│   ├── thompson2020-vcf
│   │   ├── ots34-5Mb-snpEff.vcf.gz
│   │   └── ots34-5Mb-snpEff.vcf.gz.csi
│   └── win-non-win-abs-diffs-gt0.5.rds
└── tex                                                    # Directory of files for typesetting the manuscript
    ├── .gitignore
    ├── GitLatexDiffIt.sh
    ├── abstract.tex
    ├── author-title-etc.tex
    ├── body-text-of-paper.tex
    ├── citation.bib
    ├── data-accessibility.tex
    ├── eca_molecolres.sty
    ├── images                                             # Figures/images included in manuscript.  Many (but not all) are produced when running the notebooks
    │   ├── Fall-run-ball-crop.pdf
    │   ├── Late-fall-run-ball-crop.pdf
    │   ├── Spring-run-ball-crop.pdf
    │   ├── Winter-run-ball-crop.pdf
    │   ├── ass-table-80-crop.pdf
    │   ├── banner.pdf
    │   ├── ckmr-comp-figure-crop.pdf
    │   ├── clumpak-crop.pdf
    │   ├── fpr-fnr-figure-crop.pdf
    │   ├── frh-comp-figure-crop.pdf
    │   ├── genomic-locations-plot.pdf
    │   ├── gsi_and_fst_fig-crop.pdf
    │   ├── lfar-assoc-faceted-min-0.png
    │   ├── lfar-candidates.pdf
    │   ├── major-modes-crop.pdf
    │   ├── map-crop.pdf
    │   ├── minor-modes-crop.pdf
    │   ├── months-BCF_Butte-Creek.pdf
    │   ├── months-BCS_Butte-Creek.pdf
    │   ├── months-BCkF_Blue-Creek.pdf
    │   ├── months-CHLF_Coleman-National-Fish-Hatchery.pdf
    │   ├── months-CRHS_Cole-Rivers-Hatchery.pdf
    │   ├── months-CoRHS_Cole-Rivers-Hatchery.pdf
    │   ├── months-ERF_Van-Arsdale-Fisheries-Station.pdf
    │   ├── months-FRHF_Feather-River-Hatchery.pdf
    │   ├── months-FRHS_Feather-River-Hatchery.pdf
    │   ├── months-IGHF_Iron-Gate-Hatchery.pdf
    │   ├── months-MDF_Deer-Creek.pdf
    │   ├── months-MDF_Mill-Creek.pdf
    │   ├── months-MDS_Deer-Creek.pdf
    │   ├── months-MDS_Mill-Creek.pdf
    │   ├── months-RRF_Warm-Springs-Hatchery.pdf
    │   ├── months-SJRF_San-Joaquin-River.pdf
    │   ├── months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf
    │   ├── months-SmRF_Smith-River.pdf
    │   ├── months-TRF_Trinity-River-Hatchery.pdf
    │   ├── months-TRS_Trinity-River-Hatchery.pdf
    │   ├── num-alle-barplot.pdf
    │   ├── rosa-gsi-table-crop.pdf
    │   ├── winter-v-non-winter-mh-plot_nind_ge10.pdf
    │   ├── wrap-candi.pdf
    │   └── wrap-slide-window.pdf
    ├── inputs                                              # LaTeX input for tables in manuscripts. Many (but not all) are produced when running the notebooks
    │   ├── amplicon-summary.tex
    │   ├── ascertainment-samples-supp-table.tex
    │   ├── lfar-freqs.tex
    │   ├── popgen-summary.tex
    │   ├── samples-table.tex
    │   └── wrap-freqs.tex
    ├── main-one-col.tex                                   # Main TeX document for one-column format typesetting (for submission)
    ├── main.bbl
    ├── main.tex                                           # Main TeX document for two-column format typesetting
    ├── men.bst
    ├── newcomms.tex
    ├── responses-to-referees
    │   ├── .gitignore
    │   └── responses-to-reviewers-1.Rmd
    ├── supp_data                                          # Supplemental data
    │   ├── Supp-Data-1-Amplicon-info.csv
    │   └── Supp-Data-2-pop-gen-summaries-by-locus-and-collection.csv
    ├── supplement.bbl
    ├── supplement.tex                                    # Main TeX document for typesetting the supplement
    ├── supplemental-figures.tex
    ├── supplemental-tables.tex
    └── supplemental-text.tex

```

@@

The following RMarkdown documents can be evaluated in any order.

Outputs (figures, tables, etc) from each RMarkdown document
are written to the `tex/images` or `tex/inputs` directories.  Intermediate files
written during evaluation of each RMarkdown document are written
into the `results` directories.

Thumbnails of the figures and tables generated by each RMarkdown document appear
below each notebook that generates them.

Click the thumbnail to see a full size PDF.

Additionally, at the top of each section is a link to the
compiled (HTML-version) of the RMarkdown document on GitHub Pages.

## `r make_link("001-map-of-samples.Rmd")`

Makes a map of where the samples were drawn.

[![](./docs/images_and_thumbs/figure1_thumb.png)][fig1]

## `r make_link("002-GSI_and_Fst.Rmd")`

Summaries of the samples and population genetic statistics,
as well as GSI power analysis and Fst calculation.

[![](./docs/images_and_thumbs/table1_thumb.png)][tab1]
[![](./docs/images_and_thumbs/figure2_thumb.png)][fig2]
[![](./docs/images_and_thumbs/figureS6_thumb.png)][figS6]
[![](./docs/images_and_thumbs/figureS7_thumb.png)][figS7]

[![](./docs/images_and_thumbs/supp_data2_thumb.png)][supp_data2]
[![](./docs/images_and_thumbs/table4_thumb.png)][tab4]

## `r make_link("003-ckmr-sim-evaluation.Rmd")`

Using CKMRsim to estimate false positive rates and false negative rates
of relationship inference.  Also comparisons of power for inferring relatives
when using microhaplotypes versus single SNPs from each amplicon.

[![](./docs/images_and_thumbs/figure4_thumb.png)][fig4]
[![](./docs/images_and_thumbs/figureS8_thumb.png)][figS8]
[![](./docs/images_and_thumbs/figure5_thumb.png)][fig5]

## `r make_link("004-lfar-and-wrap-allele-freqs.Rmd")`

Assessing allele frequency differences between groups of fish at the
LFAR and WRAP amplicons.

[![](./docs/images_and_thumbs/table3_thumb.png)][tab3]
[![](./docs/images_and_thumbs/tableS2_thumb.png)][tabS2]

## `r make_link("005-plotting-genome-locations-of-markers.Rmd")`

Plotting the positions of different amplicons in the genome.

[![](./docs/images_and_thumbs/figureS1_thumb.png)][figS1]

## `r make_link("006-winter-run-vs-all-cv-alle-freqs.Rmd")`

Analyzing the allele frequencies from WGS data to identify regions
that show large allele frequency differences between winter run
and everyone else.  (Finding the WRAP markers...)

[![](./docs/images_and_thumbs/figureS9_thumb.png)][figS9]
[![](./docs/images_and_thumbs/figureS10_thumb.png)][figS10]
[![](./docs/images_and_thumbs/figureS11_thumb.png)][figS11]

## `r make_link("007-whoa-on-chinook.Rmd")`

This was not included in the paper, but is an interesting analysis of
the markers suggesting that two of them might have some genotyping
issues and would be good candidates for removal in a further update
to the baseline.

[![](./docs/images_and_thumbs/whoa_thumb.png)][whoa]

## `r make_link("008-lfar-gwas-and-markers.Rmd")`

Analyzing the GWAS and allele frequency data from the LFAR search

[![](./docs/images_and_thumbs/figureS3_thumb.png)][figS3]

[fig1]: `r hdir`images_and_thumbs/figure1.pdf
[tab1]: `r hdir`images_and_thumbs/table1.pdf
[fig2]: `r hdir`images_and_thumbs/figure2.pdf
[figS6]: `r hdir`images_and_thumbs/figureS6.pdf
[figS7]: `r hdir`images_and_thumbs/figureS7.pdf
[supp_data2]: `r hdir`images_and_thumbs/Supp-Data-2-pop-gen-summaries-by-locus-and-collection.csv
[tab4]: `r hdir`images_and_thumbs/table4.pdf
[fig4]: `r hdir`images_and_thumbs/figure4.pdf
[figS8]: `r hdir`images_and_thumbs/figureS8.pdf
[fig5]: `r hdir`images_and_thumbs/figure5.pdf
[tab3]: `r hdir`images_and_thumbs/table3.pdf
[tabS2]: `r hdir`images_and_thumbs/tableS2.pdf
[figS1]: `r hdir`images_and_thumbs/figureS1.pdf
[figS9]: `r hdir`images_and_thumbs/figureS9.pdf
[figS10]: `r hdir`images_and_thumbs/figureS10.pdf
[figS11]: `r hdir`images_and_thumbs/figureS11.pdf
[whoa]: `r hdir`images_and_thumbs/whoa.pdf
[figS3]: `r hdir`images_and_thumbs/figureS3.pdf
